---
title: "Example_Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example_Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
```

## Loading Necessary Packages

```{r setup, message = FALSE}
library(tidyverse)
library(halfplus7)
```

## Saving the Data

```{r}
download_and_save <- function(file_name, url) {
  data_folder <- "../data"
  file_path <- file.path("../data", file_name)
  # See if the file already exists
  if (!file.exists("../data")) {
    dir.create("../data", recursive = TRUE)
  }
  if (!file.exists(file_path)) {
    # Download the data if it does not exist
    data <- readr::read_csv(url)
    # Save the data to the path with the desired name
    write.csv(data, file = file_path, row.names = FALSE)
  }
}

download_and_save("age_gaps.csv", 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-02-14/age_gaps.csv')
```

## Loading Dataset

```{r}
# Import dataset
age_gaps_raw <- read_csv("../data/age_gaps.csv")
```

## Cleaning Dataset 

```{r}
age_gaps_v1 <- age_gaps_raw %>%
  mutate(orientation = ifelse(character_1_gender == "man" & character_2_gender == "man" |
                               character_1_gender == "woman" & character_2_gender == "woman",
                             "LGBTQ",
                             "Straight"),
         release_decade = case_when(
           release_year %in% c(1935:1949) ~ "1935-1949",
           release_year %in% c(1950:1959) ~ "1950-1959",
           release_year %in% c(1960:1969) ~ "1960-1969",
           release_year %in% c(1970:1979) ~ "1970-1979",
           release_year %in% c(1980:1989) ~ "1980-1989",
           release_year %in% c(1990:1999) ~ "1990-1999",
           release_year %in% c(2000:2009) ~ "2000-2009",
           release_year %in% c(2010:2022) ~ "2010-2022",
           TRUE ~ NA
         ),
         actor_1_birthdate = ymd(actor_1_birthdate),
         actor_2_birthdate = ymd(actor_2_birthdate))


age_gaps_straight <- age_gaps_v1 %>%
  filter(orientation == "Straight") %>%
  mutate(man_age = ifelse(character_1_gender == "man", actor_1_age, actor_2_age),
         woman_age = ifelse(character_1_gender == "woman", actor_1_age, actor_2_age),
         man_birthdate = ifelse(character_1_gender == "man", actor_1_birthdate, actor_2_birthdate),
         woman_birthdate = ifelse(character_1_gender == "woman", actor_1_birthdate, actor_2_birthdate),
         man_name = ifelse(character_1_gender == "man", actor_1_name, actor_2_name),
         woman_name = ifelse(character_1_gender == "woman", actor_1_name, actor_2_name),         
         older_actor = case_when(
           man_birthdate < woman_birthdate ~ "man",
           woman_birthdate < man_birthdate ~ "woman",
           TRUE ~ "same birthdate"
    )
  )
```

## Visualizing Age Gaps over Time

```{r}
ggplot(age_gaps_straight, aes(x = release_year, y = age_difference, color = older_actor)) +
  geom_jitter(alpha = 0.5) +
  scale_color_manual(values = c("man" = "darkgreen", "woman" = "purple"),
    labels = c("Man", "Woman")) +
  labs(
    title = "Movie Couple Age Differences Over Time",
    x = "Year",
    y = "Age Difference",
    color = "Gender of Older \nActor in Couple:"
  ) +
  theme_minimal()
```

## Age Comparison of Men and Women in Movie Couples

```{r}
ggplot(age_gaps_straight, aes(man_age, woman_age)) +
  geom_jitter(size = 1, alpha = 0.75, color = "darkblue") +  # Add points for the scatter plot
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  theme_minimal() +
  labs(
    title = "Age Comparison of Woman and Man in Straight Movie Couples",
    subtitle = "Observations that fall on the red dashed line indicate the same age for the man and woman.",
    x = "Man's Age",
    y = "Woman's Age"
  )
```

## Creates Dataset for Straight Couples

```{r}
top_age_gaps <- age_gaps_straight %>%
  arrange(desc(age_difference)) %>%
  group_by(older_actor) %>%
  slice_head(n = 10)

long_data <- top_age_gaps %>%
  uncount(weight = 1) %>%
  pivot_longer(cols = c(man_age, woman_age), 
               names_to = "actor", 
               values_to = "age")

long_data <- long_data %>%
  mutate(actor_name = ifelse(actor == "man_age", man_name, woman_name))
```

## Visualizing Top 10 Largest Age Gaps when Man is Older

```{r}

midpoints <- long_data %>%
  group_by(movie_name) %>%
  summarize(midpoint = mean(c(actor_1_age, actor_2_age)))

long_data %>%
  filter(older_actor == "man") %>%
  ggplot(aes(x = age, y = reorder(movie_name, age_difference))) +
  geom_point(aes(color = actor)) +
  geom_segment(aes(x = age, xend = ifelse(actor == "woman_age", age + age_difference, age), 
                   y = movie_name, yend = movie_name), 
               color = "black", linetype = "solid", size = 0.5) +
  geom_text(aes(label = actor_name, y = movie_name), 
            vjust = -1, hjust = 0.5, size = 3) +
  labs(title = "Top 10 Largest Age Gaps in Straight Movie Couples \nWhen the Man is Older",
       x = "Actor Age (years)", 
       y = "Movie Title",
       color = "Gender:",
       caption = "Created using TidyTuesday data") +
  scale_color_manual(values = c("man_age" = "darkgreen", "woman_age" = "purple"),
    labels = c("Man", "Woman")) +
  theme_minimal() + 
  theme(legend.position = "top")
```

## Visualizing Top 10 Largest Age Gaps when Woman is Older

```{r}
long_data %>%
  filter(older_actor == "woman") %>%
  ggplot(aes(x = age, y = reorder(interaction(movie_name, age_difference), age_difference))) +
  geom_point(aes(color = actor)) +
  geom_segment(aes(x = age, xend = ifelse(actor == "man_age", age + age_difference, age), 
                   y = interaction(movie_name, age_difference), yend = interaction(movie_name, age_difference)), 
               color = "black", linetype = "solid", size = 0.5) +
  geom_text(aes(label = actor_name, y = interaction(movie_name, age_difference)), 
            vjust = -1, hjust = 0.5, size = 3) +
  labs(title = "Top 10 Largest Age Gaps in Straight Movie Couples \nWhen the Woman is Older",
       x = "Actor Age (years)", 
       y = "Movie Title",
       color = "Gender:",
       caption = "Created using TidyTuesday data") +
  scale_color_manual(values = c("man_age" = "darkgreen", "woman_age" = "purple"),
    labels = c("Man", "Woman")) +
  theme_minimal() + 
  theme(legend.position = "top") +
  guides(color = guide_legend(title = "Gender"))



```



### Functions Used

`dplyr` : `case_when`, `filter`, `mutate`, `slice_head`, `arrange`, `group_by`

`tidyr` : `pivot_longer`, `uncount`

`lubridate` : `ymd`

`ggplot2` : `geom_jitter`, `geom_abline`, `geom_point`, `geom_segment`, `geom_text`

`halfplus7` :



