---
title: "Example_Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example_Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
```

## Loading Necessary Packages

```{r setup, message = FALSE}
# Loads tidyverse
library(tidyverse)

# Loads halfplus7 -- the package of interest
library(halfplus7)
```

## Saving the Data

```{r}
download_and_save <- function(file_name, url) {
  data_folder <- "../data"
  file_path <- file.path("../data", file_name)
  # See if the file already exists
  if (!file.exists("../data")) {
    dir.create("../data", recursive = TRUE)
  }
  if (!file.exists(file_path)) {
    # Download the data if it does not exist
    data <- readr::read_csv(url)
    # Saves the data to the path with the desired name
    write.csv(data, file = file_path, row.names = FALSE)
  }
}

# Uses download_and_save function with age_gaps.csv from GitHub
download_and_save("age_gaps.csv", 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-02-14/age_gaps.csv')
```

## Loading Dataset

```{r}
# Imports raw age_gap dataset
age_gaps_raw <- read_csv("../data/age_gaps.csv")
```

## Cleaning Dataset 

```{r}
age_gaps_v1 <- age_gaps_raw %>%
  # creates new variable to describe sexual orientation of couple
  mutate(orientation = ifelse(character_1_gender == "man" & character_2_gender == "man" |
                               character_1_gender == "woman" & character_2_gender == "woman",
                             "LGBTQ",
                             "Straight"),
         # modifies birthdate variables to be treated as dates
         actor_1_birthdate = ymd(actor_1_birthdate),
         actor_2_birthdate = ymd(actor_2_birthdate))

age_gaps_straight <- age_gaps_v1 %>%
  # only includes straight couples
  filter(orientation == "Straight") %>%
  # create new variables to identify characteristics of actors listed by their gender
  mutate(
    man_age = case_when(
      character_1_gender == "man" ~ actor_1_age, 
      TRUE ~ actor_2_age),
    woman_age = case_when(
      character_1_gender == "woman" ~ actor_1_age, 
      TRUE ~ actor_2_age),
    man_birthdate = case_when(
      character_1_gender == "man" ~ actor_1_birthdate, 
      TRUE ~ actor_2_birthdate),
    woman_birthdate = case_when(
      character_1_gender == "woman" ~ actor_1_birthdate, 
      TRUE ~ actor_2_birthdate),
    man_name = case_when(
      character_1_gender == "man" ~ actor_1_name, 
      TRUE ~ actor_2_name),
    woman_name = case_when(
      character_1_gender == "woman" ~ actor_1_name, 
      TRUE ~ actor_2_name),  
    # identifies the gender of the older actor in the couple
    older_actor = case_when(
      man_birthdate < woman_birthdate ~ "man",
      woman_birthdate < man_birthdate ~ "woman",
      TRUE ~ "same birthdate"
    )
  )
```

## Visualizing Age Gaps over Time

```{r}
ggplot(age_gaps_straight, aes(x = release_year, y = age_difference, color = older_actor)) +
  # adds jitter for discrete data
  geom_jitter(alpha = 0.5) +
  # designates color of points by gender of older actor in couple
  scale_color_manual(values = c("man" = "darkgreen", "woman" = "purple"),
    labels = c("Man", "Woman")) +
  # creates labels for plot
  labs(
    title = "Movie Couple Age Differences Over Time",
    x = "Year",
    y = "Age Difference",
    color = "Gender of Older \nActor in Couple:"
  ) +
  theme_minimal()
```

## Age Comparison of Men and Women in Movie Couples

**Please note that this plot includes a demonstration of the check_couple function.** 

```{r}
age_gaps_straight %>%
  # demonstrates check_couple function from halfplus7 package
  mutate(match_rule = case_when(
    check_couple(man_age, woman_age) == FALSE ~ "No",
    check_couple(man_age, woman_age) == TRUE ~ "Yes")) %>%
  ggplot(aes(man_age, woman_age, color = match_rule)) +
  # adds jitter for discrete data
  geom_jitter(alpha = 0.5) +  
  # adds line at y = x to illustrate couples with the same age
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  # designates colors to identify whether couple meets 'Half-Your-Age-Plus-7' Rule
  scale_color_manual(values = c("Yes" = "blue", "No" = "red")) +
  theme_minimal() +
  # adds labels for plot
  labs(
    title = "Age Comparison of Woman and Man in Straight Movie Couples",
    subtitle = "Observations that fall on the black dashed line indicate the same age for the man and woman.",
    x = "Man's Age",
    y = "Woman's Age",
    color = "Does the Couple follow the \n'Half-Your-Age-Plus-Seven' Rule?"
  ) + 
  # moves legend to top of plot
  theme(legend.position = "top")
```

## Creates Datasets for the Straight Couples with the largest age gaps

```{r}
# creates data set with couples with largest age gaps, 10 with older man and 10 with older woman
top_age_gaps <- age_gaps_straight %>%
  arrange(desc(age_difference)) %>%
  group_by(older_actor) %>%
  slice_head(n = 10)

# modifies dataset above 
top_age_gaps_long <- top_age_gaps %>%
  uncount(weight = 1) %>%
  pivot_longer(cols = c(man_age, woman_age), 
               names_to = "actor", 
               values_to = "age") %>%
  mutate(actor_name = ifelse(actor == "man_age", man_name, woman_name),
         age_lowerlim = lower_limit(age),
         age_upperlim = upper_limit(age))
```

## Top 10 Largest Age Gaps when Man is Older

```{r}
top_age_gaps_long %>%
  filter(older_actor == "man") %>%
  ggplot(aes(x = age, y = reorder(movie_name, age_difference))) +
  geom_segment(aes(x = age, xend = ifelse(actor == "man_age", age_lowerlim, age),
                   y = movie_name, yend = movie_name), 
               color = "lightgreen", alpha = 0.25, linetype = "solid", size = 5) +
  geom_text(data = . %>% filter(actor == "man_age"),
            aes(label = paste("lower limit =", age_lowerlim), x = age_lowerlim, y = movie_name), 
            vjust = 0, hjust = 0, size = 2.5, color = "darkgreen") +
  geom_segment(aes(x = age, xend = ifelse(actor == "woman_age", age + age_difference, age), 
                   y = movie_name, yend = movie_name), 
               color = "gray", linetype = "dashed", linewidth = 0.5) +
    geom_label(aes(label = age, color = actor), show.legend = FALSE, 
             fill = "white", size = 3, hjust = 0.5, vjust = 0.5) +
    geom_text(aes(label = actor_name, y = movie_name), 
            vjust = -1.2, hjust = 0.5, size = 3) +
  labs(title = "Top 10 Largest Age Gaps in Straight Movie Couples \nWhen the Man is Older",
       subtitle = "Green shaded segments illustrate the lower age range for older actors, \ngiven the 'Half-Your-Age-Plus-7' Rule. If the younger actor's age is not \nincluded in this range that indicates that the couple would not meet the \nRule.", 
       x = "Actor Age (years)", 
       y = "Movie Title",
       color = "Gender:",
       caption = "Created using TidyTuesday data") +
  scale_color_manual(values = c("man_age" = "darkgreen", "woman_age" = "purple"),
    labels = c("Man", "Woman")) +
  theme_minimal() + 
  theme(legend.position = "top") +
  xlim(c(12, 80))
```

## Top 10 Largest Age Gaps when Woman is Older

```{r}
top_age_gaps_long %>%
  filter(older_actor == "woman") %>%
  ggplot(aes(x = age, y = reorder(interaction(movie_name, age_difference), age_difference))) +
  geom_segment(aes(x = age, xend = ifelse(actor == "man_age", age_upperlim, age),
                   y = reorder(interaction(movie_name, age_difference), age_difference), yend = reorder(interaction(movie_name, age_difference), age_difference)), 
               color = "lightgreen", alpha = 0.25, linetype = "solid", size = 5) +
  geom_text(data = . %>% filter(actor == "man_age"),
            aes(label = paste("upper limit =", age_upperlim), x = age + 2, y = reorder(interaction(movie_name, age_difference), age_difference)), 
            vjust = 0, hjust = 0, size = 2.5, color = "darkgreen") +
  geom_segment(aes(x = age, xend = ifelse(actor == "man_age", age + age_difference, age), 
                   y = interaction(movie_name, age_difference), yend = interaction(movie_name, age_difference)), 
               color = "gray", linetype = "dashed", linewidth = 0.5) +
  geom_label(aes(label = age, color = actor), show.legend = FALSE, 
             fill = "white", size = 3, hjust = 0.5, vjust = 0.5) +
  geom_text(aes(label = actor_name, y = interaction(movie_name, age_difference)), 
            vjust = -1.2, hjust = 0.5, size = 3) +
  labs(title = "Top 10 Largest Age Gaps in Straight Movie Couples \nWhen the Woman is Older",
      subtitle = "Green shaded segments illustrate the upper age range for younger actors, \ngiven the 'Half-Your-Age-Plus-7' Rule. If the younger actor's age is not \nincluded in this range that indicates that the couple would not meet the \nRule.", 
       x = "Actor Age (years)",
       x = "Actor Age (years)", 
       y = "Movie Title",
       caption = "Created using TidyTuesday data") +
  scale_color_manual(values = c("man_age" = "darkgreen", "woman_age" = "purple"),
    labels = c("Man", "Woman")) +
  theme_minimal() + 
  theme(legend.position = "top") +
  guides(color = guide_legend(title = "Gender:")) +
  xlim(c(12, 85))
```

### Functions Used

`dplyr` : `case_when`, `filter`, `mutate`, `slice_head`, `arrange`, `group_by`

`tidyr` : `pivot_longer`, `uncount`

`lubridate` : `ymd`

`ggplot2` : `geom_jitter`, `geom_abline`, `geom_point`, `geom_segment`, `geom_text`, `geom_label`

`halfplus7` : `check_couple`, `upper_limit`, `lower_limit`
